---
title: "market-segmentation.rmd"
author: "Nemo"
date: "8/16/2021"
output: pdf_document
---

```{r 4.setup, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = TRUE)

knitr::opts_chunk$set(set.seed(7))
rm(list = ls())
require(cluster)
require(ggplot2)
```

# Market Segmentation

## Data Pre-Processign

In our Datase, we observe the following:

- _X_ column is essentialy the Turk's ID. Thus, this column has been set as the row index and dropped consequently
- Mean-normalised all columns and scaled them to a 0-1 rage _(Performed Z-transformation)_ to make the measurements comparable

```{r 4.1, include=FALSE}
# clustering
# principal component analysis

NutrientH20 <- read.csv("./data/social_marketing.csv")
length(unique(NutrientH20$X)) == dim(NutrientH20)[1]
# TRUE. Thus X column is the ID

row.names(NutrientH20) <- NutrientH20$X
NutrientH20 <- NutrientH20[,-1]
columns <- colnames(NutrientH20)
NutrientH20 <- data.frame(NutrientH20)
NutrientH20 <- scale(NutrientH20)
```

## Elbow method

```{r 4.elbow}
#Elbow Method for finding the optimal number of clusters
# Compute and plot wss for k = 2 to k = 15.
k.max <- 25
wss <- sapply(1:k.max, 
              function(k){kmeans(NutrientH20, k, nstart=5,iter.max = 1000 )$tot.withinss})

plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```
The Elbow Plot above indicates a kink at _**K = 14**_. This could be a good estimate of the number of clusters we go choose.

However, This method is too _visual_ and the differences are not clearly discernible.
Thus, let us re-validate our results with the calculation of the **GAP** metric.

## Gap Index
```{r 4.GAP}
nH20_gap <- clusGap(NutrientH20, FUN = kmeans, nstart = 5,
                       K.max = 20, B = 3, iter.max = 1000)
plot(nH20_gap)
nH20_gap$Tab

find_peaks <- function (x, m = 3){
     shape <- diff(sign(diff(x, na.pad = FALSE)))
     pks <- sapply(which(shape < 0), FUN = function(i){
        z <- i - m + 1
        z <- ifelse(z > 0, z, 1)
        w <- i + m + 1
        w <- ifelse(w < length(x), w, length(x))
        if(all(x[c(z : i, (i + 2) : w)] <= x[i + 1])) return(i + 1) else return(numeric(0))
    })
     pks <- unlist(pks)
     pks
}

k.max <- find_peaks(nH20_gap$Tab[,"gap"], m = 1)[1]

cat("\n\nOptimum cluster size is:",paste(k.max),"\n")
```

From the above plots and output statistic, we see that the local optima of the Cluster Gap plot is achieved at _**K = 14**_ clusters. Thus, we go ahead with dividing the consumer base in 14 clusters for targeted marketing.

## Running K-Means with 14 clusters
```{r 4.kmeans, message=TRUE}
# Run k-means with 215 clusters and 25 starts
clust1 = kmeans(NutrientH20, k.max, nstart=25, iter.max = 1000)


# What are the clusters?
for(clusterID in c(1: k.max))
{
  data_cluster <- NutrientH20[(clust1$cluster == clusterID),]
  s_mean <- apply(data_cluster,2,mean)
  s_sd <- apply(data_cluster,2,sd)
  
  cat("\n\nMetrics for Cluster #",paste(clusterID),":")
  for(colID in columns){
    cat("\n\t",paste(colID),"center:",paste(clust1$centers[clusterID,colID]),",sd:",paste(s_sd[colID]),",95% Confidence Interval: (", paste(clust1$center[clusterID,colID]-1.96*s_sd[colID]),",", paste(clust1$center[clusterID,colID]+1.96*s_sd[colID]),")")
  }
  cat("\n")
}
```

```{r 4.plot, message=TRUE, warning=FALSE}
# A few plots with cluster membership shown
# qplot is in the ggplot2 library
qplot(NutrientH20[,"religion"], xlab = "Religion",
      NutrientH20[,"sports_fandom"], y_lab = "Sports Fandom",
      color=factor(clust1$cluster))
qplot(NutrientH20[,"politics"], xlab = "Politics",
      NutrientH20[,"news"], y_lab = "News",
      color=factor(clust1$cluster))
# Compare versus within-cluster average distances from the first run
cat("\nCluster Whithinness:",paste(clust1$withinss))
cat("\nTotal Whithinness:",paste(clust1$tot.withinss))
cat("\nCluster Betweenness:",paste(clust1$betweenss))
```
